Viscous Erosion in Porous Media

### 2018-03-04 docs/methods/erosion_methods.tex
NM

As a final validation, we now aim to verify that our method converges with second-order accuracy in time. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes the body to vanish, and was chosen so that the body will have eroded substantially but is not so small that accuracy is compromised. We assess error by measuring the difference between simulations of successive resolutions (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time. At each stage we halve the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy. In the table, we also show the CPU time of the simulation run on a single processor (though the algorithm is easily parallelized), showing a linear increase in computational time.

@@

We therefore impose the condition of a fixed pressure difference by controlling the incoming flow-rate, $\umax$, at each time-step. This is achieved by first computing the Stokes flow with $\umax = 1$ prescribed, then measuring the pressure at a location upstream ($x = -2$) and downstream ($x=2$) of the eroding bodies. The vertically-averaged upstream and downstream pressures are given by
\begin{align}
p_{-} = \frac{1}{2} \int_{-1}^{1} p(-2,y) \, dy \, , \qquad
p_{+} = \frac{1}{2} \int_{-1}^{1} p(+2,y) \, dy \, .
\end{align}
We then rescale the computed density function by the pressure difference
\begin{align}
\eeta \leftarrow \frac{\eeta}{(p_{-} - p_{+})} \, .
\end{align}
This rescaling maintains a {\em fixed} pressure difference. Because the Stokes equations are linear, such a rescaling is permissible and does not alter the accuracy of our method. Furthermore, rescaling by a different factor at each time-step is also permissible because the Stokes equations are memoryless (i.e.~there is no time derivative present in the PDEs).

>>>>

As a final validation, we now aim to verify that our method converges with second-order accuracy in time. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes the body to vanish, and was selected so that the body will have eroded substantially but is not so small that accuracy is compromised. We assess error by measuring the difference between simulations of successive resolutions (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time. At each stage we halve the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy. In the table, we also show the CPU time of the simulation run on a single processor (though the algorithm is easily parallelized), showing a linear increase in computational time.

@@

We therefore impose the condition of a fixed pressure difference by dynamically controlling the incoming flow-rate, $\umax$. This is achieved by first computing the Stokes flow with $\umax = 1$ prescribed, then measuring the pressure at a location upstream ($x = -2$) and downstream ($x=2$) of the eroding bodies. The vertically-averaged upstream and downstream pressures are given by
\begin{align}
p_{-} = \frac{1}{2} \int_{-1}^{1} p(-2,y) \, dy \, , \qquad
p_{+} = \frac{1}{2} \int_{-1}^{1} p(+2,y) \, dy \, .
\end{align}
We then rescale the computed density function by the pressure difference
\begin{align}
\eeta \leftarrow \frac{\eeta}{(p_{-} - p_{+})} \, .
\end{align}
This rescaling maintains a {\em fixed} pressure difference between the upstream ($x=-2$) and downstream ($x=2$) positions. Because the Stokes equations are linear, such a rescaling is permissible and does not alter the accuracy of our method. Furthermore, {\em dynamic} rescaling is also permissible because the Stokes equations are memoryless (i.e.~there is no time derivative present in the PDEs).
